---
title: "downsize"
author: "Will Landau"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{downsize}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

With the `downsize` package, you can toggle the test and production versions of your workflow with the flip of a `TRUE/FALSE` global option. This is helpful when your workflow takes a long time to run, but you want to test it quickly and [unit testing](http://kbroman.org/pkg_primer/pages/tests.html) is too reductionist to cover everything.

# Toggle objects

Say you want to analyze a large dataset.

```{r}
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4))
```

But for the sake of time, you want to test and debug your code on a smaller dataset. In your code, select your dataset with a call to `downsize()`.

```{r}
library(downsize)
my_data <- downsize(big_data) # downsize(big = big_data)
```

Above, `my_data` becomes `big_data` if `getOption("downsize")` is `FALSE` or `NULL` (default) and `head(big_data)` if `getOption("downsize")` is `TRUE`. 
You can toggle the global option `downsize` with calls to `scale_up()` and `scale_down()`, and you can override the option with `downsize(..., downsize = L)`, where `L` is `TRUE` or `FALSE`. Check if you are in test mode or production mode with the `scaling()` function.

```{r}
scale_down()
scaling() # getOption("downsize") is TRUE
my_data <- downsize(big_data) # downsize(big = big_data)
identical(my_data, head(big_data))
scale_up()
scaling() # getOption("downsize") is FALSE
my_data <- downsize(big_data) # downsize(big = big_data)
identical(my_data, big_data)
```

Equivalently, you can provide a replacement for `big_data` using argument `small` in `downsize()`.

```{r}
small_data <- data.frame(x = runif(16), y = runif(16))
scale_down()
scaling() # getOption("downsize") is TRUE
my_data <- downsize(big_data, small_data) # downsize(big = big_data, small = small_data)
identical(my_data, small_data)
```

If you set `small` yourself, be sure that subsequent code can accept both `small` and `big`. For example, if `small` is a data frame and `big` is a matrix, your code may work fine in test mode and break in production mode. In addition, `downsize()` will warn you if `small` is identical to or bigger in memory than `big` (disable with `downsize(..., warn = FALSE`)). See the next section for a safer alternative.

# Toggle subsetting

The command `my_data <- downsize(big = big_data)` is equivalent to `my_data <- downsize(big = big_data, nrow = 6)`. There are multiple ways to subset argument `big` in `downsize()` when it is time to scale down. As in the following examples, be sure that `small` is set to `NULL` (default).

```{r}
scale_down()
downsize(1:10, length = 2)
m <- matrix(1:36, ncol = 6)
downsize(m, ncol = 2)
downsize(m, nrow = 2)
downsize(m, dim = c(2, 2))
downsize(data.frame(x = 1:10, y = 1:10), nrow = 5)
x = array(0, dim = c(10, 100, 2, 300, 12))
dim(x)
my_array <- downsize(x, dim = rep(3, 5))
dim(my_array)
my_array <- downsize(x, dim = c(1, 4))
dim(my_array)
my_array <- downsize(x, ncol = 1)
dim(my_array)
```

Set `random` to `TRUE` to take a random subset of your data rather than just the first few rows or columns.

```{r}
set.seed(6)
downsize(m, ncol = 2, random = T)
```

# Toggle code blocks

You can interchange entire blocks of code based on the scaling of the workload.

```{r}
scale_down()
downsize(big = {a = 1; a + 10}, small = {a = 1; a + 1})
scale_up()
downsize(big = {a = 1; a + 10}, small = {a = 1; a + 1})
```

Variables set in code blocks are available after calls to `downsize()`.

```{r}
scale_down()
tmp <- downsize(
  big = {
    x = "long code"
    y = 1000
  }, 
  small = {
    x = "short code"
    y = 3.14
  })
x == "short code" & y == 3.14
scale_up()
tmp <- downsize(
  big = {
    x = "long code"
    y = 1000
  }, 
  small = {
    x = "short code"
    y = 3.14
  })
x == "long code" & y == 1000
```

The `downsize()` function checks the values of `big` and `small`, so any  code blocks passed as arguments must return non-`NULL` values. To avoid warnings, the return value of `small` should be different from and no larger in memory than that of `big` (or just call `downsize(..., warn = FALSE)`). If subsetter arguments such as `dim` and `length` are used, they apply to the return value of `big` rather than any variables defined in the code block.

# Troubleshooting

Please refer to [TROUBLESHOOTING.md](https://github.com/wlandau/downsize/blob/master/TROUBLESHOOTING.md) on the [GitHub page](https://github.com/wlandau/downsize) for instructions.
