---
title: "downsize"
author: "Will Landau"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{downsize}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

If you have time-consuming workflow, it may be prudent to try a downsized
dry run first. If you intersperse your code with enough calls to `downsize::ds()`, 
it is trivially easy to scale the workload up or down. 
That way, your entire workflow becomes a quick unit test with as little change as possible.

```{r}
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4))
```

But for the sake of time, you want to test and debug your code on a smaller dataset.

```{r}
small_data <- head(big_data)
```

In your code, define your dataset with a call to `ds()`.

```{r}
library(downsize)
my_data <- ds(big_data, small_data)
```

The `ds()` function executes `my_data <- big_data` if the `getOption("downsize")` is `FALSE` or `NULL` (default) and `my_data <- small_data` if `TRUE`. You can toggle the global option `downsize` with a call to `scale_up()` or `downsize()` (same as `scale_down()`), and you can check the status with `scaling()`.

```{r}
downsize() # same as scale_down()
scaling()
my_data <- ds(big_data, small_data)
all(my_data == small_data)
scale_up()
scaling()
my_data <- ds(big_data, small_data)
all(my_data == big_data)
```

In this case, `my_data <- ds(big_data, small_data)` is equivalent to `my_data <- ds(big_data, nrow = 6)`. You can override the global option `downsize` with `ds(..., downsize = L)`, where `L` is `TRUE` or `FALSE`.

For more examples, run the following lines in a new R session. Then, enter `scale_down()` and run those lines again to see what changes.

```{r, eval = F}
library(downsize)
ds("Leave me alone when the downsize option is FALSE.", 
  "Return me when downsize option is TRUE.")
ds(1:10, length = 2)
m <- matrix(1:36, ncol = 6)
ds(m, ncol = 2)
ds(m, ncol = 2, random = T)
ds(m, nrow = 2)
ds(m, dim = c(2, 2))
ds(data.frame(x = 1:10, y = 1:10), nrow = 5)
dim(ds(array(0, dim = c(10, 100, 2, 300, 12)), dim = rep(3, 5)))
```

