---
title: "Package tutorial: downsize"
author: "Will Landau"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package tutorial: downsize}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

If you have time-consuming workflow, it may be prudent to try a downsized
dry run first. That way, you can quickly test your entire workflow from start to finish before deploying your next serious job. 
If you intersperse your code with calls to the `ds()` function in the `downsize` package, it is trivially easy to scale down or up. Say you want to analyze a large dataset.

```{r}
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4))
```

But for the sake of time, you want to test and debug your code on a smaller dataset.

```{r}
small_data <- head(big_data)
```

In your code, define your dataset with a call to `ds()`.

```{r}
library(downsize)
my_data <- ds(big_data, small_data)
```

The `ds()` function executes `my_data <- big_data` if the `getOption("downsize")` is `FALSE` (default) and `my_data <- small_data` otherwise. You can toggle the global option `downsize` with a call to `scale_up()` or `scale_down()` (or write `ds(..., downsize = FALSE)`), and you can check the status with `scaling()`.

```{r}
scale_down()
scaling()
my_data <- ds(big_data, small_data)
all(my_data == small_data)
scale_up()
scaling()
my_data <- ds(big_data, small_data)
all(my_data == big_data)
```

In this case, `my_data <- ds(big_data, small_data)` is equivalent to `my_data <- ds(big_data, nrow = 6)`.

For more examples, run the following lines in a new R session. Then, enter `scale_down()` and run those lines again to see what changes.

```{r, eval = F}
library(downsize)
ds("Leave me alone when the downsize option is FALSE.", 
  "Return me when downsize option is TRUE.")
ds(1:10, length = 2)
m <- matrix(1:36, ncol = 6)
ds(m, ncol = 2)
ds(m, ncol = 2, random = T)
ds(m, nrow = 2)
ds(m, dim = c(2, 2))
ds(data.frame(x = 1:10, y = 1:10), nrow = 5)
dim(ds(array(0, dim = c(10, 100, 2, 300, 12)), dim = rep(3, 5)))
```

