---
title: "downsize"
author: "Will Landau"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{downsize}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

With the `downsize` package, you can toggle the test and production versions of your workflow with the flip of a `TRUE/FALSE` global option. This is helpful when your workflow takes a long time to run, you want to test it quickly, and [unit testing](https://www.r-bloggers.com/unit-testing-with-r/) is too reductionist to cover everything.

# Basic downsizing

Say you want to analyze a large dataset.

```{r, eval = F}
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4))
```

But for the sake of time, you want to test and debug your code on a smaller dataset. In your code, select your dataset with a call to `downsize()`.

```{r, eval = F}
my_data <- downsize(big_data) # downsize(big = big_data)
```

Above, `my_data` becomes `big_data` if `getOption("downsize")` is `FALSE` or `NULL` (default). If `getOption("downsize")` is `TRUE`, `big_data` becomes `head(big_data)`. You can toggle the global option `downsize` with calls to `production_mode()` and `test_mode()`, and you can override the option with `downsize(..., downsize = L)`, where `L` is `TRUE` or `FALSE`. Check if the workflow is in test or production mode with the `my_mode()` function.

# Example with test and production modes

Here is an example script in test mode.

```{r, eval = F}
library(downsize)
test_mode() # scales the workflow appropriately
my_mode() # shows if the workflow is in test or production mode
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4)) # always large
my_data <- downsize(big_data) # either large or small
nrow(my_data) # responds to test_mode() and production_mode()
# ...more code, time-consuming if my_data is large...
```

To scale up the workflow up to production mode, replace `test_mode()` with `production_mode()` and **leave everything else exactly the same**.

```{r, eval = F}
library(downsize)
production_mode() # scales the workflow appropriately
my_mode() # shows if the workflow is in test or production mode
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4)) # always large
my_data <- downsize(big_data) # either large or small
nrow(my_data) # responds to test_mode() and production_mode()
# ...more code, time-consuming if my_data is large...
```

An ideal workflow has multiple calls to `downsize()` that are configured all at once with a single call to `test_mode()` or `production_mode()` at the very beginning. Thus, tedium and human error are avoided, and the test is a close approximation to the original task at hand.

# Provide your own test data

You can provide a replacement for `big_data` using argument `small` in `downsize()`.

```{r}
library(downsize)
big_data <- data.frame(x = rnorm(1e4), y = rnorm(1e4))
small_data <- data.frame(x = runif(16), y = runif(16))
test_mode()
my_mode() # getOption("downsize") is TRUE
my_data <- downsize(big_data, small_data) # downsize(big = big_data, small = small_data)
identical(my_data, small_data)
```

If you set `small` yourself, be sure that subsequent code can accept both `small` and `big`. For example, if `small` is a data frame and `big` is a matrix, your code may work fine in test mode and break in production mode. In addition, `downsize()` will warn you if `small` is identical to or bigger in memory than `big` (disable with `downsize(..., warn = FALSE`)). To be safer, use the subsetting capabilities of the `downsize()` function.

# Subsetting

The command `my_data <- downsize(big = big_data)` is equivalent to `my_data <- downsize(big = big_data, nrow = 6)`. There are multiple ways to subset argument `big` in `downsize()` when it is time to scale down to test mode. As in the following examples, be sure that `small` is set to `NULL` (default). Otherwise, subsetter arguments such as `dim`, `length`, `nrow`, and `ncol` will be ignored.

```{r}
test_mode()
downsize(1:10, length = 2)
m <- matrix(1:36, ncol = 6)
downsize(m, ncol = 2)
downsize(m, nrow = 2)
downsize(m, dim = c(2, 2))
downsize(data.frame(x = 1:10, y = 1:10), nrow = 5)
x = array(0, dim = c(10, 100, 2, 300, 12))
dim(x)
my_array <- downsize(x, dim = rep(3, 5))
dim(my_array)
my_array <- downsize(x, dim = c(1, 4))
dim(my_array)
my_array <- downsize(x, ncol = 1)
dim(my_array)
```

Set `random` to `TRUE` to take a random subset of your data rather than just the first few rows or columns.

```{r}
set.seed(6)
downsize(m, ncol = 2, random = T)
```

# Interchange code blocks

You can interchange entire blocks of code based on the scaling/mode of the workload.

```{r}
test_mode()
downsize(big = {a = 1; a + 10}, small = {a = 1; a + 1})
production_mode()
downsize(big = {a = 1; a + 10}, small = {a = 1; a + 1})
```

Variables set in code blocks are available after calls to `downsize()`.

```{r}
test_mode()
tmp <- downsize(
  big = {
    x = "long code"
    y = 1000
  }, 
  small = {
    x = "short code"
    y = 3.14
  })
x == "short code" & y == 3.14
production_mode()
tmp <- downsize(
  big = {
    x = "long code"
    y = 1000
  }, 
  small = {
    x = "short code"
    y = 3.14
  })
x == "long code" & y == 1000
```

# Help and troubleshooting

Use the `help_downsize()` function to obtain a collection of helpful links. For troubleshooting, please refer to [TROUBLESHOOTING.md](https://github.com/wlandau/downsize/blob/master/TROUBLESHOOTING.md) on the [GitHub page](https://github.com/wlandau/downsize) for instructions.
